In-silico molecular enrichment and clearance of the human intracranial space

### 2025-06-10 paper/main.tex
remove hornkjol ref
Updates in Methods

For an extended description of the mathematical model and simulation
scenarios, see the Supplementary information S1.

@@

\subsection*{CSF flow in the SAS and ventricular system}

We model CSF as an incompressible Newtonian fluid at low Reynolds and
Womersley numbers via the Stokes equations with viscosity $\mu$, and
account for the convective contribution of the flow induced by CSF
production in the choroid plexus. Specifically, we compute the
velocity $\bm u_{\rm CSF}$ and associated pressure $p_{\rm CSF}$ in
the SAS and ventricular system induced by a constant production at a
rate of $u_{\rm in}$ across the lateral ventricle walls and with
efflux across the upper, outer (arachnoid) boundary with efflux
resistance $R_0$; and then set $\bm u = \bm u_{\rm CSF}$ in
$\Omega_{\rm CSF}$ in~\eqref{eq:multi_transport}
(\Cref{tab:parameters}, \Cref{fig:csf}A--E, see also
the Supplementary information (S1)). We do not model bulk flow within the brain
parenchyma, except within the perivascular network segments extending
below the pial surface, and thus set $\bm u = 0$ in $\Omega_{\rm PAR}$
in~\eqref{eq:multi_transport}.

\subsection*{Dispersion in the CSF spaces}

The pulsatile flow of CSF in the SAS and ventricular system,
associated with the cardiac and respiratory cycles, substantially
enhances molecular diffusion through
dispersion~\cite{taylor1953dispersion, watson1983diffusion,
  asgari2016glymphatic, sharp2019dispersion, ray2021quantitative,
  troyetsky2021dispersion}. To account for these dispersive effects,
while bridging from the pulsatile flow time scale of seconds to a
molecular transport time scale of hours, we compute cardiac and
respiratory dispersion factors $R_c$ and $R_r$, both
spatially-varying, in $\Omega_{\rm CSF}$ (\Cref{fig:csf}A) by
combining computational estimates of the respective peak CSF pressure
gradients with theoretical estimates for shear-augmented (Taylor)
dispersion \cite{taylor1953dispersion, watson1983diffusion,
  sharp2019dispersion} (see Supplementary information S1.3 for a
complete description). Combining these effects, we then set $R = R_c +
R_r$ and define the effective diffusion coefficient $D = (1 + R)
D^{\rm Gad}$ in $\Omega_{\rm CSF}$ in~\eqref{eq:multi_transport} in
the baseline model.

\subsection*{Perivascular fluid flow induced by CSF pressure differences}

To estimate the contribution also to perivascular fluid flow from CSF
production, we impose the fluid pressure $p_{\rm CSF}$ induced by CSF
production, computed in the SAS and ventricular system, at the end
nodes of the periarterial network located within the SAS. For the
network end nodes located within the parenchyma, we compute a harmonic
extension of the CSF pressure field (by solving a Laplace equation in
$\Omega_{\rm PAR}$ with $p_{\rm CSF}$ as the boundary value), and set
the corresponding value at these end nodes. We then numerically solve
a system of hydraulic network equations to compute $\hat{u}^p_a$
defined over $\Lambda_a$; i.e., the pressure-difference induced CSF
flow velocity in the periarterial network (see
also Supplementary information S1.1.4). We repeat this procedure to
compute a corresponding velocity $\hat{u}^p_v$ in the perivenous
network $\Lambda_v$.

@@

      Molecular efflux rate & $\beta_{\rm exit}$ & $8\cdot 10^{-7}$ & m/s & \cite{hornkjol2022csf} \\

@@

      CSF outflow resistance & $R_0$ & $10^{4}$  & Pa$\cdot$s/m & \cite{hornkjol2022csf} \\ 

>>>>

This section gives a condensed summary of the in-silico modelling
framework. For a complete description of the mathematical models and
simulation scenarios, we refer to the Supplementary information S1.

@@

\subsection*{Steady flow in the CSF spaces induced by CSF production}

We model CSF as an incompressible Newtonian fluid at low Reynolds and
Womersley numbers via the Stokes equations with viscosity $\mu$ (see
\Cref{tab:parameters} for parameter values). To account for steady
flow induced by CSF production, we compute the CSF velocity field $\bm
u_{\rm CSF}$ and associated pressure field $p_{\rm CSF}$ in the SAS
and ventricular system $\Omega_{\rm CSF}$ that result from a steady
CSF production at a rate of $u_{\rm in}$ across the lateral ventricle
walls (\Cref{fig:csf}A--E). We allow for CSF efflux across the upper,
outer (arachnoid) boundary with efflux resistance $R_0$.

%then set $\bm u = \bm u_{\rm CSF}$ in $\Omega_{\rm CSF}$
%in~\eqref{eq:multi_transport}

%% We do not model bulk flow within the brain parenchyma, except within
%% the perivascular network segments extending below the pial surface,
%% and thus set $\bm u = 0$ in $\Omega_{\rm PAR}$
%% in~\eqref{eq:multi_transport}.

\subsection*{Perivascular fluid flow induced by CSF production}

To estimate the contribution also to perivascular fluid flow from CSF
production, we impose the fluid pressure $p_{\rm CSF}$ induced by CSF
production, computed in the SAS and ventricular system, at the end
nodes of the periarterial network. For the end nodes located within
the SAS ($v \in \Omega_{\rm CSF}$), the value $p_{\rm CSF}(v)$ is used
directly, while for the end nodes located within the parenchyma ($v \in
\Omega_{\rm PAR}$), we compute and apply a harmonic extension
$E(p_{\rm CSF})(v)$. We then numerically solve a system of hydraulic
network equations to compute $\hat{u}^p_a$ defined over $\Lambda_a$;
i.e., the CSF production-induced velocity field in the periarterial
network. We refer to Supplementary information S1.1.4 \emph{Steady
flow in perivascular networks induced by pressure differences} for
details. We repeat this procedure to compute a corresponding velocity
field $\hat{u}^p_v$ in the perivenous network $\Lambda_v$.

\subsection*{Dispersion in the CSF spaces induced by pulsatile CSF flow}

The pulsatile flow of CSF in the SAS and ventricular system,
associated with the cardiac and respiratory cycles, substantially
enhances molecular diffusion through
dispersion~\cite{taylor1953dispersion, watson1983diffusion,
  asgari2016glymphatic, sharp2019dispersion, ray2021quantitative,
  troyetsky2021dispersion}. To account for these dispersive effects,
while bridging from the pulsatile flow time scale of seconds to a
molecular transport time scale of hours, we estimate cardiac- and
respiratory dispersion factors ($R_c$ and $R_r$) in $\Omega_{\rm CSF}$
(\Cref{fig:csf}G, I). To compute these spatially-varying factors, we
combine computational estimates of peak CSF pressure gradients with
theoretical estimates for shear-augmented (Taylor) dispersion
\cite{taylor1953dispersion, watson1983diffusion, sharp2019dispersion}
(see Supplementary information S1.3 Estimating dispersion factors from
pulsatile CSF flow for a complete description). Combining the cardiac
and respiratory effects, we set $R = R_c + R_r$ in $\Omega_{\rm CSF}$.

%and define the effective diffusion coefficient $D = (1 + R) D^{\rm
%Gad}$ in $\Omega_{\rm CSF}$ in~\eqref{eq:multi_transport} in the
%baseline model.

@@

      Molecular efflux rate & $\beta_{\rm exit}$ & $8\cdot 10^{-7}$ & m/s & $\ast$ \\

@@

      CSF outflow resistance & $R_0$ & $10^{4}$  & Pa$\cdot$s/m & $\ast$ \\ 
